---
title: "p8130_hw6_ps3194"
author: "Pangsibo Shen"
date: "12/3/2020"
output: html_document
---

```{r}
library(tidyverse)
```
### Problem 1

```{r}
#load the data
pat_satisfaction = read_csv("./data/PatSatisfaction.csv") %>%
  janitor::clean_names()
```

##### 1)

Create a correlation matrix for all variables and interpret your findings. Focus on the correlation values between each predictor and the outcome of interest.

```{r}
round(cor(pat_satisfaction),3)

# Scatter plot matrix for all variables
pairs(pat_satisfaction)
```

##### 2)

Fit a multiple regression model including all three predictors and test whether at least one of these variables is significant. State the hypotheses, test-statistic, decision rule and conclusion.


$$
\begin{split}
H_0: \beta_1 = \beta_2 =\beta_3 =0\\
H_1: the \ least \ one \ \beta \ is \ not \ zero\\
F = \frac{MSR}{MSE}=\frac{3040.155}{101.1629}=30.05208 \\
F(1-\alpha;p,n-p-1) = 2.816466
\end{split}
$$

```{r}
fit_1 = lm(safisfaction~age + severity + anxiety, data = pat_satisfaction)
anova(fit_1)
MSR = sum(anova(fit_1)[1:3, 'Sum Sq'])/3
MSE = anova(fit_1)['Residuals', 'Mean Sq']
F_stats = MSR/MSE
qf(0.95,3,44)
summary(fit_1)
```
For the F statistics, we got `r F_stats` which is greater than the F critical value (`r qf(0.95,3,44)`). Hence we reject the null hypothesis and conclude that at significance level of 0.05, there are some predictors are nonzero.








